### 2022-10-07 Proposal/proposal.tex
add proposal latex file with basic template

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}

%% preamble
\usepackage[margin = 1in]{geometry}

%% metadata
\title{Proposal}
\author{James Horbury\\
    University of Connecticut
}
\date{October 10, 2022}

\begin{document}
\maketitle

\section*{Introduction}
\addcontentsline{toc}{section}{Introduction}
\label{sec:intro}

\section*{Specific Aims}
\addcontentsline{toc}{section}{Specific Aims}
\label{sec:spec}

\section*{Data Description}
\addcontentsline{toc}{section}{Data Description}
\label{sec:data}

\section*{Research Design/Methods/Schedule}
\addcontentsline{toc}{section}{Research Design/Methods/Schedule}
\label{sec:res}

\section*{Discussion}
\addcontentsline{toc}{section}{Discussion}
\label{sec:disc}

\section*{Conclusion}
\addcontentsline{toc}{section}{Conclusion}
\label{sec:conc}

\bibliographystyle{chicago}
\bibliography{citations}

\end{document}